#Steps for prepping Access database for R code that aggregates hourly to daily data:
#  1) Copy/Paste record from Access into Excel
#  2) Remove spaces within cells
#  3) Date:
#     Format cells -> custom: Yyyy-mm-dd
#  4) Time:
#     Format cells -> time: 24-hr time
#  5) Save as .CSV file

# Raw data file format:
#####################################
         site       date  time   temp
1 FinneyBrook 2013-08-16 18:00 13.690
2 FinneyBrook 2013-08-16 18:30 13.618
3 FinneyBrook 2013-08-16 19:00 13.570
4 FinneyBrook 2013-08-16 19:30 13.546
5 FinneyBrook 2013-08-16 20:00 13.522
6 FinneyBrook 2013-08-16 20:30 13.473
#####################################

```{r Set directories and specify some parameters}
rm(list=ls())

library(plyr)

setwd('F:/KPONEIL/SourceData/streamTemperature/MA')

# What units are the data in? C or F? (or K)
deg <- 'C'

# Number of decimals to round the output to
numDecimals <- 3

# What agency did the data come from
Agency <- 'CTDEEP'
```

```{r Read in data and merge}


# Reclassify date column
temp$date <- as.Date(temp$date)

# If necessary, convert from Fahrenheit to Celsius
if (deg == 'F') { temp$temp <- (temp$temp - 32)*(5/9) } 

```

```{r Brief QA/QC}

# Make sure min and max temps are reasonable (0 to ~ 25-30)
print(range(temp$temp, na.rm = T))

# Make all negative stream temperatures equal to 0.
temp$temp[which(temp$temp < 0)] <- 0

```
      
```{r Get min/mean/max}      

# Min/mean/max
finalData <- ddply( temp, .(site, date), summarise, meanTemp = round( mean(temp,na.rm=T), digits = numDecimals),
                                                    minTemp  = round( min (temp,na.rm=T), digits = numDecimals), 
                                                    maxTemp  = round( max (temp,na.rm=T), digits = numDecimals))

finalData <- replace(finalData, is.na(finalData), NA)



# Fill missing dates with NA values in between min and max dates:
# ===============================================================
sites <- unique(finalData$site)
tf <- finalData

for ( i in 1:length(sites) ){
  
  expData <- tf[tf$site == sites[i],]
  
  startDate <- min(expData$date)
  endDate   <- max(expData$date)  
  
  Record <- data.frame(date = seq(from=as.Date(startDate),to=as.Date(endDate),by="day"), site = expData$site[1])
  
  newRecord <- merge(Record, expData, by = c('date', 'site'), all.x = T, all.y = F, sort = F)

  if(i == 1) { finalData <- newRecord} else ( finalData <- rbind(finalData, newRecord))
  
}

finalData$year <- as.numeric(strftime(finalData$date, '%Y'))
finalData$dOY <- as.numeric(strftime(finalData$date, '%j'))

finalData <- merge(finalData, meta, by = 'site', all.x = T, sort = F)

finalData$agency <- Agency

finalData$AgencyID <- finalData$site

finalData$site <- paste0(Agency, '_', finalData$site)


save(finalData, file = 'WestbrookStreamTempDate.RData')


```





```{r Pair air temperature}

load('F:/KPONEIL/SourceData/climate/UMASS (Westbrook)/WestbrookMetData.RData')

air <- wbMetData[,c('date', 'meanAirTemp')]
names(air) <- c('date', 'airTemp')

head(finalData)

merge(finalData)


```

#Loop through all sites and fill gaps with NAs
Sites <- unique(finalData$site)
tempFrame <- finalData

for ( i in 1:length(Sites) ){
  
  expData <- tempFrame[tempFrame$site == Sites[i],]
  
  start.date <- min(expData$date)
  end.date   <- max(expData$date)  
  
  Record <- data.frame(seq(from=as.Date(start.date),to=as.Date(end.date),by="day"))
  names(Record) <- "date"
  Record$year <- as.numeric(strftime(Record$date, '%Y'))
  Record$dOY <- as.numeric(strftime(Record$date, '%j'))
  
  newRecord <- merge(Record, expData, by = c('date', 'dOY', 'year'), all.x = T, all.y = F, sort = F)
  
  #Fill in blanks:
  newRecord$site       <- expData$site[1]
  newRecord$AgencyID   <- expData$AgencyID[1]
  newRecord$agency     <- expData$agency[1]
  #newRecord$Latitude   <- expData$Latitude[1]
  #newRecord$Longitude  <- expData$Longitude[1]
  
  if(i == 1) { finalData <- newRecord} else ( finalData <- rbind(finalData, newRecord))
  
}


masterData <- merge(finalData, sites, by = 'site', all.x = T, sort = F)

save(masterData,file="dailyStreamTempVTFWS.Rdata")
